The Cluster Difference Imaging Photometric Survey (CDIPS). I. Light Curves \& Planet Candidates in Open Clusters from TESS Sectors 6 \& 7

### 2019-06-13 paper/ms.tex
update

\shortauthors{Bouma et al.}
\shorttitle{CDIPS I: Method, Sectors 6 \& 7}

@@

% BEGIN NUMBERS
\newcommand{\numberpcs}{10\ }
\newcommand{\numberlcs}{42{,}424\ }
\newcommand{\numberclusters}{424\ }

@@

\begin{abstract}
  % Context
  The Transiting Exoplanet Survey Satellite (TESS) has started to
  photometrically monitor almost every star cluster in
  the solar neighborhood.
  % Aims
  As part of the CDIPS project, we have been processing the TESS
  full-frame images to make high-precision light curves stars that
  could be members of stellar associations, moving groups, and open clusters.
  %
  Our eventual hope is to discover giant transiting planets, and probe
  the exoplanet size and separation distributions for planets younger
  than 1 Gyr.
  %
  For this study, we made \numberlcs light curves of candidate young
  stars in \numberclusters clusters.
  Each light curve represents between 20 and 25 days of observations
  of a star brighter than $G_{Rp}=16$, sampled at 30 minute cadence.
  % Methods
  We describe in detail the image subtraction and time-series analysis 
  techniques we used to create the light curves, which
  have noise properties in agreement with theoretical expectations.
  % Results
  We then highlight the \numberpcs planet candidates we identified in
  an initial search, describe the vetting procedures that they have
  survived, and emphasize the value that additional
  spectroscopic, photometric, and imaging observations will have in
  verify that these candidates are in fact planets.
  %
  We also comment on the possible use of the light curve sample for
  studies of stellar rotation evolution, and binary eccentricity
  damping.
  %
  The light curves and vetting are available at
  \url{archive.stsci.edu/prepds/cdips}.
\end{abstract}

@@

Many evolutionary processes are expected to occur over the lives of
stars and exoplanets; it would be interesting to statistically observe
these changes as functions of time.
However, since dwarf stars do not appreciably change in size or luminosity
until evolving off the main sequence, creative approaches must be used to
infer ages \citep{soderblom_ages_2010}.
In the study of exoplanets, one approach beginning to bear fruit
is to search for planets in open clusters and moving groups.
The ages of these stellar ensembles are inferred through isochrone
analysis, and are typically most constrained by the location of the
main-sequence turnoff and/or the red giant branch (CITE IF TRUE).
Once calibrated, gyrochronology can also be used to verify
the age of a stellar ensemble \citep[{\it
e.g.},][]{barnes_color-period_2015,meibom_spin-down_2015,curtis_tess_2019}.
The ages of planets discovered around these cluster stars are then
assumed to equal those of the star.

The first planet detections in open clusters
came from radial velocity measurements of evolved stars in the Hyades
\citep{Sato_et_al_2007}, and in NGC~2423 \citep{lovis_mayor_2007}.
Subsequent spectroscopic searches in Praesepe led to the detection of the first
hot Jupiters in a cluster \citep{Quinn_et_al_2012}, and also the first
multi-planet system \citep{Malavota_et_al_2016}.
There were also additional detections reported in the Taurus
star-forming region (CITE V830 Tau b; Donati et al. 2016, and Cl Tau
b; Johns-Krull et al. 2016, Flagg et al. 2019).
An additional radial velocity survey was conducted by Brucalassi et al
(2017, CITE), following 88 stars in M~67 over 7 years.  They
reported the detections of three hot Jupiters around main sequence
stars, and one giant planet around an evolved star.

The transit method was slow to catch up, despite deep searches of the
M~37 open cluster by \citet{hartman_MMT_IV_2009}, as well as
multi-year searches of nine young clusters in the solar neighborhood
by the Monitor project
\citep{aigrain_monitor_2007,irwin_monitordata_2007,miller_monitor_2008}.
While these studies led to interesting measurements of stellar
rotation \citep{hartman_rotation_2009,Irwin_NGC2516_2007} as well as
the discovery of pre-main-sequence (PMS) eclipsing binaries (EBs)
\citep{irwin_monitor_PMSEB_2007}, no planet in open clusters were
uncovered.  Transit searches in globular clusters were similarly met
with null results \citep{gilliland_lack_2000,
weldrake_searching_2006}.

Kepler \citep{borucki_kepler_2010} was the first machine to observe a
sufficient number of stars for long enough and with sufficient
precision to detect transiting planets in open clusters:  Kepler-66b
and 67b, in the gigayear-old NGC6811 \citep{Meibom_et_al_2013}.
Though a broken reaction wheel ended the prime Kepler mission, the
repurposed K2 mission \citep{howell_k2_2014} switched between
fields along the ecliptic every quarter-year, and was able to observe
far more clusters and young stars.
In particular, K2's detections through its surveys of the Pleiades,
Hyades, Praesepe, and Upper Sco were a major inspiration for the
present work.

K2-25b, discovered in the Hyades, had a radius ($3.43\,R_\oplus$) that
was larger than older Kepler planets at similar insolation levels
\citep{Mann_K2_25_2016}.  This inflated radius is perhaps a hint that
the planet is contracting, or losing its atmosphere (e.g., CITE).
The super-Neptune sized K2-33b ($P\approx 5.4\,{\rm d}$), found in
Upper Sco, showed that at least some close-in planets must form within
about 10 Myr, excluding long-timescale migration as an explanation for
its origin \citep{Mann_K2_33b_2016,David_et_al_2017}.
Further discoveries included six transiting planets in Praesepe (CITE
Obermeier, Mann 2016 or 2017), a long-period candidate super-Earth in
the Hyades (CITE Vanderburg 2017), K2-136b in the Hyades (CITE
Ciardi+2018, Livingston+2018), the transiting multi-planet system
K2-264bc in Praesepe (CITE Rizzuto+2018, Livingston+2019), as well as
a Jupiter-sized planet orbiting every 24 days around V1298 Tau (CITE
David+2019).
A final related detection was that of three small planets transiting
K2-233 (CITE David 2018).  Though the star was not in a cluster, its
age was determined to be $360_{-140}^{+490}\,{\rm Myr}$ based on
rotation, activity, and kinematic indicators.

Clearly, a significant amount of effort has been and continues to be
devoted towards the discovery and characterization of planets around
young stars.
It would be nice to begin occurrence rate testing, with a machinery
similar to that described by Rizzuto 2017 (CITE).
However, due to the geometric rarity of transits, many tens of
thousands of stars with known ages must be monitored at high precision
for such studies to be statistically robust.

@@

We have begun to produce time-series photometry from the TESS images
for stars in ``clusters''~--~a collective term we adopt for open
clusters, moving groups, stellar associations, and star-forming
regions.
We are also including some stars that we suspect are young due to
combined photometric and astrometric indicators (see
\S~\ref{sec:starselection}).
Our eventual hope is to discover giant transiting planets, and probe
the exoplanet size and separation distributions for planets younger
than 1 Gyr.

%FIXME: this paragraph needs improvement.
Unfortunately, there are formidable obstacles to deriving precise
photometry from the {\it TESS} images because of the relatively poor
angular resolution.  Almost all of the clusters are within 10 degrees
of the Galactic plane (see FIGURE~1), where the problems with crowding
and complex backgrounds are so severe that the {\it TESS}\, Science
Team has neglected that portion of the sky in their planet simulations
\citep{Sullivan_et_al_2015}. Similarly, the {\it TESS}\, Candidate
Target List deprioritizes all objects close to the galactic
plane~\citep{stassun_TIC_2018,stassun_TIC8_2019}, which includes 90\%
of all star clusters.  The large pixel size and the high stellar
surface density will make aperture photometry unreliable.  Yet,
aperture photometry is the basis of the official {\it TESS} data
reduction pipeline and almost all reduction methods that have been
applied to {\it Kepler} and {\it K2} data.
We have therefore adopted the difference imaging technique (also
dubbed ``image subtraction''), because of its benefits in modelling
complex backgrouds in crowded regions.

In what follows, \S~\ref{sec:method} presents BLAH, and
\S~\ref{sec:results} describes BLAH BLAH.
\S~\ref{sec:discussion} discusses, and \S~\ref{sec:conclusion}
concludes.

@@

First, to identify periodic transit-like signals, we used 
\citet{hippke_TLS_2019}'s transit least-squares (TLS) tool.
The algorithm is the same as the canonical box least-squares
\citep{kovacs_box-fitting_2002}, except in place of a box template, a transit 
template is used  for a marginal improvement of the detection efficiency.
In addition, the search grids in \texttt{tls} 
\footnote{\url{https://github.com/hippke/tls}} are slightly more 
efficient than in most BLS implementations, since \texttt{tls} uses the 
cubic-in-frequency sampling advocated by \citet{ofir_optimizing_2014}, 
rather than standard linear-in-frequency sampling.
During the period search, we rejected $6\,{\rm hours}$ at the
beginning and end of each spacecraft orbit, to mitigate the presence
of correlated red noise in the results.
This shrinks the data volume by about 5\%, but also lowers the number
of systematic false positives in subsequent vetting.
Our grids typically consisted of about 25 different durations, and
about 2500 periods between 0.5 and 11 days. 
We also ran a generalized Lomb-Scargle periodogram (CITE), simply to 
have additional periodicity information.

>>>>

\shortauthors{Bouma et al.}
\shorttitle{CDIPS I}

@@

% BEGIN NUMBERS
\newcommand{\numberpcs}{42\ }
\newcommand{\numberlcs}{99{,}999\ }
\newcommand{\numberclusters}{XXX\ }

@@

\begin{abstract}
  % Context
  The Transiting Exoplanet Survey Satellite (TESS) has started to
  photometrically monitor almost every star cluster in
  the solar neighborhood.
  % Aims
  As part of the CDIPS project, we have been processing the TESS
  full-frame images to make light curves of stars that
  could be members of stellar associations, moving groups, and open clusters.
  %
  Our eventual hope is to discover giant transiting planets, and probe
  the exoplanet size and separation distributions for planets younger
  than 1 Gyr.
  %
  For this study, we made \numberlcs light curves of candidate young
  stars in \numberclusters clusters.
  Each light curve represents between 20 and 25 days of observations
  of a star brighter than $G_{Rp}=16$, sampled at 30 minute cadence.
  % Methods
  We describe in detail the image subtraction and time-series analysis 
  techniques we used to create the light curves, which
  have noise properties in agreement with theoretical expectations.
  % Results
  We then highlight the \numberpcs planet candidates we identified in
  an initial search, describe the vetting procedures that they have
  survived, and emphasize the value that additional
  spectroscopic, photometric, and imaging observations will have in
  verify that these candidates are in fact planets.
  %
  We also comment on the possible use of the light curve sample for
  studies of stellar rotation evolution, and binary eccentricity
  damping.
  %
  The light curves and vetting reports are available at
  \url{archive.stsci.edu/prepds/cdips}.
\end{abstract}

@@

Many evolutionary processes are expected to occur over the lives of
stars and exoplanets; it would be interesting to statistically observe
these changes as functions of time.
However, since most dwarf star properties do not appreciably change
until the stars evolve off the main sequence, creative approaches must
be used to infer stellar ages \citep{soderblom_ages_2010}.
In the study of exoplanets, one approach beginning to bear fruit
is to search for planets in open clusters and moving groups.
The ages of these stellar ensembles are inferred through isochrone
analyses, and are typically most constrained by the location of the
main-sequence turnoff and/or the red giant branch (CITE IF TRUE).
Once calibrated, gyrochronology can also be used to verify
the age of a stellar ensemble \citep[{\it
e.g.},][]{barnes_color-period_2015,meibom_spin-down_2015,curtis_tess_2019}.
The ages of planets discovered around these cluster stars are then
assumed to equal those of the star.

The first planets in open clusters were detected through radial
velocity (RV) measurements of evolved stars in the Hyades
\citep{Sato_et_al_2007}, and in NGC~2423 \citep{lovis_mayor_2007}.
Subsequent spectroscopic searches in Praesepe led to the detection of
the first hot Jupiters in a cluster \citep{Quinn_et_al_2012}, and also
the first multi-planet system \citep{Malavolta_et_al_2016}.
The youngest RV detections have come from surveys of T Tauri stars in the
Taurus star-forming region, which showed signals of the hot Jupiter
\citep[V830~Tau~b][]{donati_hj_2016} and of a giant planet on a 9-day orbit
\citep[CI~Tau~b][]{johns-krull_candidate_2016,flagg_co_2019}.
A final RV survey of note was conducted by
\citet{brucalassi_search_2017}, who followed 88 stars in M~67 over 7
years.  They reported detections of three hot Jupiters around main
sequence stars, and of one giant planet around an evolved star.

The transit method was slow to catch up, despite deep searches of
M~37 by \citet{hartman_MMT_IV_2009}, as well as
searches of nine young clusters in the solar neighborhood
by the Monitor project
\citep{aigrain_monitor_2007,irwin_monitordata_2007,miller_monitor_2008}.
While these studies led to valuable measurements of stellar
rotation \citep{hartman_rotation_2009,Irwin_NGC2516_2007} as well as
the discovery of pre-main-sequence (PMS) eclipsing binaries (EBs)
\citep{irwin_monitor_PMSEB_2007}, no planet in open clusters were
uncovered.  Transit searches in globular clusters were similarly met
with null results \citep{gilliland_lack_2000,
weldrake_searching_2006}.

Kepler \citep{borucki_kepler_2010} was the first project to observe a
sufficient number of stars for long enough and with sufficient
precision to detect transiting planets in open clusters:  Kepler-66b
and 67b, in the gigayear-old NGC~6811 \citep{Meibom_et_al_2013}.
Though a broken reaction wheel ended the prime Kepler mission, the
repurposed K2 mission \citep{howell_k2_2014} switched between
fields along the ecliptic every quarter-year, and was able to observe
far more clusters and young stars.
In particular, K2's detections through its surveys of the Pleiades,
Hyades, Praesepe, and Upper Sco were a major inspiration for the
present work.

K2-25b, discovered in the Hyades, had a radius ($3.43\,R_\oplus$) that was
larger than older Kepler planets at similar insolation levels
\citep{Mann_K2_25_2016}.  This inflated radius is perhaps a hint that the
planet is contracting, or losing its atmosphere (e.g., CITE).  The
super-Neptune sized K2-33b ($P\approx 5.4\,{\rm d}$), found in Upper Sco,
showed that at least some close-in planets must form within about 10 Myr,
excluding long-timescale migration as an explanation for its origin
\citep{Mann_K2_33b_2016,David_et_al_2017}.  Further discoveries included six
transiting planets in Praesepe \citep{obermeier_k2_2016,Mann_et_al_2017}, a
long-period candidate super-Earth in the Hyades
\citep{vanderburg_zeitVII_2018}, the three-planet K2-136 system in the Hyades
\citep{ciardi_k2-136_2018,livingston_three_2018,mann_ZEITVI_2018}, and also the
two-planet system K2-264 in Praesepe
\citep{rizzuto_zeitVIII_2018,livingston_k2-264_2019}.  A final recent
K2 detection is of a Jupiter-sized planet orbiting every 24 days around V1298
Tau \citep{david_v1298_2019}.

Clearly, a significant amount of effort has been and continues to be
devoted towards the discovery and characterization of planets around
young stars.
Sometimes, cluster membership may not even be necessary, if a star's
youth can be inferred based on a combination of rotation, activity,
and kinematic indicators \citep[{e.g.}, K2-233][]{david_k2-233_2018}.
To investigate demographics in time, the detection sensitivities and
reliabilities of these searchers need to be calibrated, likely with a
machinery similar to that outlined by \citet{rizzuto_zeitV_2017}.
A separate requirement for population studies to be statistically
robust, imposed by the geometric rarity of transits, is that many tens
of thousands of stars with known ages must be monitored at high
precision.

@@

One of the major challenges in deriving precise photometry from the
TESS images is the relatively poor angular resolution ($\approx
21''$).  Almost all clusters are within 10 degrees of the Galactic
plane (see FIGURE~1). The problems with crowding and complex
backgrounds are so severe near the galactic plane that the {\it
TESS}\, Candidate Target List has deprioritized all objects at
galactic latitudes less than $\approx
15^\circ$~\citep{stassun_TIC_2018,stassun_TIC8_2019}, which includes
90\% of all star clusters\footnote{TICv8 updates this to $10^\circ$,
but for the first year of TESS observations, the number used when
selecting target stars was $15^\circ$.}.
This is sensible, given that the large pixel size and the high stellar
surface density will make aperture photometry unreliable.  However, it
also means that a large number of stars in clusters that could yield
planet candidates may go unprocessed by the official {\it TESS} data
reduction pipeline.

We have therefore opted to produce time-series photometry from the
TESS images for stars in ``clusters''~--~a collective term we adopt
for open clusters, moving groups, stellar associations, and
star-forming regions.
We are also including some stars that we suspect are young due to
combined photometric and astrometric indicators.
We are using the difference imaging technique because of its benefits
in modelling complex backgrouds in crowded regions.
Our eventual hope is to discover giant transiting planets, and probe
the exoplanet size and separation distributions for planets younger
than 1 Gyr.

%FIXME improve
In what follows, \S~\ref{sec:starselection} describes the process of
selecting stars.  \S~\ref{sec:method} presents the photometric and
image processing methods, and
\S~\ref{sec:results} describes results.
\S~\ref{sec:discussion} discusses, and \S~\ref{sec:conclusion}
concludes.

@@

First, to identify periodic transit-like signals, we used 
\citet{hippke_TLS_2019}'s transit least-squares (TLS) tool.
The algorithm is the same as the canonical box least-squares
\citep{kovacs_box-fitting_2002}, except in place of a box template, a transit 
template is used  for a marginal improvement of the detection efficiency.
In addition, the search grids in \texttt{tls} 
\footnote{\url{https://github.com/hippke/tls}} are slightly more 
efficient than in most BLS implementations, since \texttt{tls} uses the 
cubic-in-frequency sampling advocated by \citet{ofir_optimizing_2014}, 
rather than standard linear-in-frequency sampling.
During the period search, we rejected $6\,{\rm hours}$ at the
beginning and end of each spacecraft orbit, to mitigate the presence
of correlated red noise in the results.
This shrinks the data volume by about 5\%, but also lowers the number
of systematic false positives in subsequent vetting.
Our grids typically consisted of about 30 different durations, and
about 5000 periods between 0.5 and 21 days. 
We also ran a generalized Lomb-Scargle periodogram (CITE), simply to 
have additional periodicity information.
